---
title: "Streamlined mRNA DE Lung tissue"
author: "Matthias Hulliger"
date: "5/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load multiplot function
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```
#Loading phenotype data
```{r}
library(DESeq2)
colData<-read.table("colData7.txt",header = T)
colData$Extraction_batch<-as.factor(colData$Extraction_batch)
colData$gelding<-as.factor(colData$gelding)
colData$symptomes<-relevel(colData$symptomes,"none","Asymptomatic","Symptomatic")
colData
str(colData)
```


#Load count data
```{r,fig.width=10,fig.height=6}
DESeq2Table <- DESeqDataSetFromHTSeqCount(sampleTable = colData,directory= "/Users/matthias/R_DE_mirna_analysis/lung/mrna/10112017_gene_based_analysis/counting",design = ~ gelding+Extraction_batch+symptomes)
print(DESeq2Table)
summary(DESeq2Table)
colnames(DESeq2Table)
dds<-DESeq2Table
 design(dds)
 summary(dds)
 dim(dds)
# prefiltering
 idx<-rowMeans(counts(dds)>0)>=0.9
 dds<-dds[idx,]
 dim(dds)
 dds<-DESeq(dds)
 resS<-results(dds,contrast=c("symptomes","Symptomatic","none"),alpha=0.05)
 summary(resS)
 sigS<-subset(resS,resS$padj<0.05)
 summary(sigS)
  head(sigS[order(sigS$padj),])
  sigSord<-sigS[order(abs(sigS$log2FoldChange),decreasing=T),]
  sigSordpvalue<-sigS[order(abs(sigS$pvalue),decreasing=F),]
  head(sigSord,n=11)
 rownames(sigSord)

 write.table(rownames(sigSord),"DEgenes_SvC.txt",quote = F)
 
 resA<-results(dds,contrast=c("symptomes","Asymptomatic","none"),alpha=0.05)
 summary(resA)
 sigA<-subset(resA,resA$padj<0.05)
 summary(sigA)
 head(sigA[order(sigA$padj),])
 sigAord<-sigA[order(abs(sigA$log2FoldChange),decreasing=T),]
 head(sigAord,n=11)
 rownames(sigAord)
 write.table(rownames(sigAord),"DEgenes_AvC.txt",quote = F)
 
 resSvA<-results(dds,contrast=c("symptomes","Symptomatic","Asymptomatic"),alpha=0.05)
 summary(resSvA)
 sigSvA<-subset(resSvA,resSvA$padj<0.05)
summary(sigSvA)
 sigSvAord<-sigSvA[order(abs(sigSvA$log2FoldChange),decreasing=T),]
 head(sigSvAord,n=11)
 rownames(sigSvAord)
 write.table(rownames(sigSvAord),"DEgenes_SvA.txt",quote = F)
 SvAup<-rownames(subset(sigSvAord,sigSvAord$log2FoldChange>0))
  SvAdown<-rownames(subset(sigSvAord,sigSvAord$log2FoldChange<0))
   Sup<-rownames(subset(sigSord,sigSord$log2FoldChange>0))
   Sdown<-rownames(subset(sigSord,sigSord$log2FoldChange<0))
   Adown<-rownames(subset(sigAord,sigAord$log2FoldChange<0))
   Aup<-rownames(subset(sigAord,sigAord$log2FoldChange>0))
  SvAup
  write.table(SvAup,"SvAup.txt",quote=F)
  write.table(SvAdown,"SvAdown.txt",quote=F)
  write.table(Sup,"Sup.txt",quote=F)
  write.table(Sdown,"Sdown.txt",quote=F)
  write.table(Aup,"Aup.txt",quote=F)
  write.table(Adown,"Adown.txt",quote=F)
  sigS["FCER1A",]
  sigS["MS4A2",]
  resSvA["MS4A2",]
  resSvA["FCER1A",]

 A<-subset(resA,resA$padj<=0.05)
A<-A[order(A$padj),]
  head(A,n=11)
  write.table(A,"AvC-DE_mrna.csv")
   SvAr<-subset(resSvA,resSvA$padj<=0.05)
SvAr<-SvAr[order(SvAr$padj),]
  head(SvAr,n=11)
  write.table(SvAr,"DEgenes_SvA.csv")
#  caluclate median & mean log 2 fold change:
    dim(sigSord)
  S_up<-head(subset(sigSord,sigSord$log2FoldChange>0))
 S_down<-head(subset(sigSord,sigSord$log2FoldChange<0))
     dim(sigA)
  A_up<-head(subset(sigA,sigA$log2FoldChange>0))
 A_down<-head(subset(sigA,sigA$log2FoldChange<0))
  SvA_up<-head(subset(sigSvA,sigSvA$log2FoldChange>0))
 SvA_down<-head(subset(sigSvA,sigSvA$log2FoldChange<0))
 mean(A_up$log2FoldChange)
sd(A_up$log2FoldChange)
 mean(A_down$log2FoldChange)
sd(A_down$log2FoldChange)
 mean(S_up$log2FoldChange)
sd(S_up$log2FoldChange)
 mean(S_down$log2FoldChange)
sd(S_down$log2FoldChange)
  sd(A_down$log2FoldChange)
 mean(SvA_up$log2FoldChange)
sd(SvA_up$log2FoldChange)
 mean(SvA_down$log2FoldChange)
sd(SvA_down$log2FoldChange)
  
 #plotLogFold<-function(sig,x,y,z){
  #produce a horizontal barplot with the log2 fold changes of the top down-
   #and upregulated genes/miRNAs. Input parametersA: sig : filtered results table by deseq2 (padj<0.05). n : number of n top up-and down regulated genes to plot. y and z are the plot relations according to the x axes (how plotting space is distributed according to the scale of the x axes).
 S_up<-head(subset(sigSord,sigSord$log2FoldChange>0),n=20)
 S_down<-head(subset(sigSord,sigSord$log2FoldChange<0),n=20)
 #S_up
 #S_down
 library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
 S_up<-as.data.frame(S_up)
 #S_up
 S_up<-S_up[order(S_up$log2FoldChange,decreasing=T),]
 #S_up
S_up$padj<-format(S_up$padj,scientific=T,digits=2)
 S_down<-as.data.frame(S_down)
 S_down<-S_down[order(abs(S_down$log2FoldChange),decreasing=T),]
 S_down$padj<-format(S_down$padj,scientific=T,digits=2)
 
p1<-ggplot(S_down, aes(x = reorder(row.names(S_down),-S_down$log2FoldChange), S_down$log2FoldChange)) +
        geom_bar(stat = "identity") + theme_bw(base_size=19)+labs(y="Log 2 Fold Change",x="Downregulated genes")+ 
        coord_flip(ylim=c(0,-6.6)) + theme(axis.text=element_text(size=19,color="black"),axis.title.y=element_blank(),plot.title = element_text(size = 15, face = "bold",hjust = 0.5))+
        geom_text(aes(label = S_down$padj), nudge_y = -0.9, color = "black")+ ggtitle("Top downregulated genes")+ scale_y_continuous(breaks=c(0,-2,-4,-6),expand =c(0,0))+ geom_text(aes(label = ""), nudge_y = -0.2 )

p2<-ggplot(S_up, aes(x = reorder(row.names(S_up),S_up$log2FoldChange), S_up$log2FoldChange)) + theme_bw(base_size = 19)+
        geom_bar(stat = "identity") + scale_x_discrete(position = "top",expand = c(0,0))+  labs(y="Log 2 Fold Change",x="Upregulated genes")+ 
        coord_flip(ylim=c(0,7.8)) + theme(axis.text=element_text(size=19,color="black"),axis.title.y=element_blank(),plot.title = element_text(size = 15, face = "bold",hjust = 0.5))+
        geom_text(aes(label = S_up$padj), nudge_y = +0.8, color = "black")+ ggtitle("Top upregulated genes")+scale_y_continuous(expand = c(0,0))+ geom_text(aes(label = ""), nudge_y = +0.2 )
        library(cowplot)

  plot_grid(p1, p2,align="h",rel_widths = c(1,1.05))
  r<-plot_grid(p1, p2,align="h",rel_widths = c(1.05,1))
  ggsave("EvClogPlot.tiff",r,device="tiff",dpi=300)

  #plot for asymptomatic vs control with top 15 genes
  
  S_up<-head(subset(sigAord,sigAord$log2FoldChange>0),n=20)
 S_down<-head(subset(sigAord,sigAord$log2FoldChange<0),n=20)
 S_up<-as.data.frame(S_up)
 #S_up
 S_up<-S_up[order(S_up$log2FoldChange,decreasing=T),]
 #S_up
S_up$padj<-format(S_up$padj,scientific=T,digits=2)
 S_down<-as.data.frame(S_down)
 S_down<-S_down[order(abs(S_down$log2FoldChange),decreasing=T),]
 S_down$padj<-format(S_down$padj,scientific=T,digits=2)
 
p1<-ggplot(S_down, aes(x = reorder(row.names(S_down),-S_down$log2FoldChange), S_down$log2FoldChange)) +
        geom_bar(stat = "identity") + theme_bw(base_size=19)+labs(y="Log 2 Fold Change",x="Downregulated genes") + theme(axis.text=element_text(size=19,color="black"),axis.title.y=element_blank(),plot.title = element_text(size = 15, face = "bold",hjust = 0.5))+
        geom_text(aes(label = S_down$padj), nudge_y = -0.7, color = "black")+ ggtitle("Top downregulated genes")+ scale_y_continuous(expand =c(0,0))+ geom_text(aes(label = ""), nudge_y = +1 )+
        coord_flip(ylim=c(0,-5.3))

p2<-ggplot(S_up, aes(x = reorder(row.names(S_up),S_up$log2FoldChange), S_up$log2FoldChange)) + theme_bw(base_size = 19)+
        geom_bar(stat = "identity") + scale_x_discrete(position = "top",expand = c(0,0))+  labs(y="Log 2 Fold Change",x="Upregulated genes")+ 
        coord_flip() + theme(axis.text=element_text(size=19,color="black"),axis.title.y=element_blank(),plot.title = element_text(size = 15, face = "bold",hjust = 0.5))+
        geom_text(aes(label = S_up$padj), nudge_y = +0.5, color = "black")+ ggtitle("Top upregulated genes") +geom_text(aes(label = ""), nudge_y = +1.2 )+ylim(0,4.5)+scale_y_continuous(expand = c(0,0))
        library(cowplot)

  plot_grid(p1, p2,align="h",rel_widths = c(1.05,1))
  q<-plot_grid(p1, p2,align="h",rel_widths = c(1.05,1))
 ggsave("RvClogPlot.tiff",q,device="tiff",dpi=300)
  #plot for symptomatic vs asymptomatic  with top 20 genes
  
  S_up<-head(subset(sigSvAord,sigSvAord$log2FoldChange>0),n=20)
 S_down<-head(subset(sigSvAord,sigSvAord$log2FoldChange<0),n=20)
 S_up<-as.data.frame(S_up)
 #S_up
 S_up<-S_up[order(S_up$log2FoldChange,decreasing=T),]
 #S_up
S_up$padj<-format(S_up$padj,scientific=T,digits=2)
 S_down<-as.data.frame(S_down)
 S_down<-S_down[order(abs(S_down$log2FoldChange),decreasing=T),]
 S_down$padj<-format(S_down$padj,scientific=T,digits=2)
 
p1<-ggplot(S_down, aes(x = reorder(row.names(S_down),-S_down$log2FoldChange), S_down$log2FoldChange)) +
        geom_bar(stat = "identity") + theme_bw(base_size=18)+labs(y="Log 2 Fold Change",x="Downregulated genes")+theme(axis.text=element_text(size=18,color="black"),axis.title.y=element_blank(),plot.title = element_text(size = 13, face = "bold"))+
        geom_text(aes(label = S_down$padj), nudge_y = -0.9, color = "black")+ ggtitle("Top downregulated genes")+ scale_y_continuous(expand =c(0,0),breaks=seq(0,-6,-2))+ geom_text(aes(label = ""), nudge_y = +1 )+coord_flip(ylim=c(0,-6))

p2<-ggplot(S_up, aes(x = reorder(row.names(S_up),S_up$log2FoldChange), S_up$log2FoldChange)) + theme_bw(base_size = 18)+
        geom_bar(stat = "identity") + scale_x_discrete(position = "top",expand = c(0,0))+  labs(y="Log 2 Fold Change",x="Upregulated genes")+ 
        coord_flip(ylim=c(0,11.2)) + theme(axis.text=element_text(size=18,color="black"),axis.title.y=element_blank(),plot.title = element_text(size = 13, face = "bold",hjust = 0.5))+
        geom_text(aes(label = S_up$padj), nudge_y = +0.9, color = "black")+ ggtitle("Top upregulated genes") +geom_text(aes(label = ""), nudge_y = +1 )+scale_y_continuous(expand = c(0,0),breaks = seq(0,11,2))
        library(cowplot)

  plot_grid(p1, p2,align="h",rel_widths = c(1,1.3))
  p<-plot_grid(p1, p2,align="h",rel_widths = c(1,1.3))
 ggsave("EvRlogPlot.tiff",p,device="tiff",dpi=300)

 
sessionInfo()
dim(sigS)
head(sigS)
head(rownames(sigS))
write.table(rownames(sigS),"DEG_SvC_ECab3.txt")
dim(sigA)
head(sigA)
head(rownames(sigA))
write.table(rownames(sigA),"DEG_AvC_ECab3.txt")
dim(sigSvA)
head(sigSvA)
head(rownames(sigSvA))
write.table(rownames(sigSvA),"DEG_SVA_ECab3.txt")
mRNA_norm_counts<-counts(dds,normalized=TRUE)
str(mRNA_norm_counts)
write.table(mRNA_norm_counts,"mRNA_norm_counts.txt")



```

#plot top 10 up and down in each group:
name<-"Boxplots_counts_DEG"
ending<-".pdf"
library("ggpubr")
library("ggsignif")
for( mirna in rownames(head(sigSordpvalue,n=20))){ #for all sign. DE. microRNAs do
filename<-paste(name,mirna,ending,sep="")
   mirna_counts<-plotCounts(dds, gene=mirna, intgroup="symptomes",returnData = TRUE)
   padj<-resSvA[mirna,]$padj
   log2fc<-resSvA[mirna,]$log2FoldChange
   log2fc<-round(log2fc,2)
    if(padj>0.001){
      padj<-round(padj,3)} else{
        padj<-format(padj,scientific=TRUE,digits = 3)
      }
 p<-ggerrorplot(mirna_counts, x = "symptomes", y = "count", 
               desc_stat = "mean_sd", color = "red",
               add = "jitter", add.params = list(color = "black",size=6),title = paste("",mirna,"\n (log2FC = ",log2fc,")"),xlab = "Condition",ylab="DESeq2 normalized read counts",size=1)+theme(plot.title = element_text(hjust = 0.5,size=33,face="bold"),axis.title=element_text(size=25,face="bold"),axis.text=element_text(size=16,colour="black",face="bold"),panel.border = element_rect(size=1.5,fill=NA))+geom_signif(comparisons = list(c("Symptomatic","Asymptomatic")),map_signif_level=FALSE,annotations=paste("P = ",padj),textsize=5.5)+ scale_x_discrete(labels=c("Control","Remission","Exacerbation"))
#print(mirna_counts)
print(p)}
#ggsave(filename,p)}
```


```{r,fig.width=8,fig.height=8}
vsd <- vst(dds, blind=FALSE)
mat<-assay(vsd)
project.pca <- prcomp(t(mat))
summary(project.pca)
project.pca.proportionvariances <- ((project.pca$sdev^2) / (sum(project.pca$sdev^2)))*100
#Plots scatter plot for PC 1 and 2
plot(project.pca$x, type="n", main="Principal components analysis plot; miRNA expression",xlab=paste("PC1, ", round(project.pca.proportionvariances[1], 2), "%"), ylab=paste("PC2, ", round(project.pca.proportionvariances[2], 2), "%"),cex=1.3,cex.lab=1.3)
points(project.pca$x, col=c("#000000","#009E73","#CC79A7"), pch=16, cex=3.5)
legend("topright", legend = levels(colData$symptomes), pch=16,title="Condition",col=c("#000000","#009E73","#CC79A7"),cex=1.5)



library(limma)
vsd <- varianceStabilizingTransformation(dds)
mat<-assay(vsd)
project.pca <- prcomp(t(mat))
summary(project.pca)
project.pca.proportionvariances <- ((project.pca$sdev^2) / (sum(project.pca$sdev^2)))*100

barplot(project.pca.proportionvariances, cex.names=1, xlab=paste("Principal component (PC), 1-", length(project.pca$sdev)), ylab="Proportion of variation (%)", main="Scree plot", ylim=c(0,100))

#Plots scatter plot for PC 1 and 2
plot(project.pca$x, type="n", main="Principal components analysis plot", xlab=paste("PC1, ", round(project.pca.proportionvariances[1], 2), "%"), ylab=paste("PC2, ", round(project.pca.proportionvariances[2], 2), "%"))
points(project.pca$x, col=colData$symptomes, pch=16, cex=2.5)
head(project.pca$x)
legend("topleft",legend=colData$symptomes,
       col=c("red", "blue"), lty=1:2, cex=0.8)
PCAscores<-project.pca$x
PCAcolors <- c("black","blue","magenta")[as.integer(colData$symptomes)]
tiff(file="PCA_mrna.tiff",
width=7, height=7, units="in", res=100)
p<-plot(PCAscores[,1:2],  # x and y data
     pch=21,           # point shape
     col=PCAcolors,    # point border color
     bg=PCAcolors,     # point color
     cex=2.5,          # point size
     main="PCA plot mRNA expression",    # title of plot
     xlab=paste("PC1, ", round(project.pca.proportionvariances[1], 2), "%"),  ylab=paste("PC2, ", round(project.pca.proportionvariances[2], 2), "%"),
     cex.lab=1.5
)
legend("topright",                                # position of legend
       legend=c("Controls","Remission","Exacerbation"),                       # legend display
       pch=21,                                    # point shape
       pt.bg= c("black","blue","magenta"),    # point colors
       pt.cex=1.5,                                # point size
       col =  c("black","blue","magenta")  )  # point border color
dev.off()

```
```{r,fig.width=6.5,fig.height=7.2}
library(ggrepel)
genes<-as.data.frame(resS)
resS$gene<-rownames(resS)
genes$Significant <- ifelse(genes$padj < 0.05, "FDR < 0.05", "Not Sig")
p<-ggplot(genes, aes(x = log2FoldChange, y = -log10(pvalue))) + geom_point(aes(color = Significant,size=10)) + scale_color_manual(values = c("red", "grey")) + theme_bw(base_size = 11) + theme(legend.position = "bottom",axis.title = element_text(size = 20)) 
p <- p + labs(title = "Volcano Plot exacerbation vs. controls")+theme(text = element_text(size = 20))
p
ggsave("volcano_SvC.tiff",p,device="tiff",dpi=300)


genes<-as.data.frame(resSvA)
resSvA$gene<-rownames(resSvA)
genes$Significant <- ifelse(genes$padj < 0.05, "FDR < 0.05", "Not Sig")
p<-ggplot(genes, aes(x = log2FoldChange, y = -log10(pvalue))) + geom_point(aes(color = Significant,size=10)) + scale_color_manual(values = c("red", "grey")) + theme_bw(base_size = 11) + theme(legend.position = "bottom",axis.title = element_text(size = 20)) 
p <- p + labs(title = "Volcano Plot exacerbation vs. remission")+theme(text = element_text(size = 20))
p
ggsave("volcano_SvA.tiff",p,device="tiff",dpi=300)

genes<-as.data.frame(resA)
resA$gene<-rownames(resA)
genes$Significant <- ifelse(genes$padj < 0.05, "FDR < 0.05", "Not Sig")
p<-ggplot(genes, aes(x = log2FoldChange, y = -log10(pvalue))) + geom_point(aes(color = Significant,size=10)) + scale_color_manual(values = c("red", "grey")) + theme_bw(base_size = 11) + theme(legend.position = "bottom",axis.title = element_text(size = 20))
p <- p + ggtitle("Volcano Plot remission vs. cotrols")+theme(text = element_text(size = 20))
p
ggsave("volcano_AvC.tiff",p,device="tiff",dpi=300)

library(dplyr)
 resS$gene<-NULL
 head(resS)
#MA-plots

p<-ggmaplot((as.data.frame(resS)), main = expression("MA-Plot exacerbation vs. controls"),
   fdr = 0.05,fc=0.0, size = 4, top = 0,
   palette = c("#B31B21", "#1465AC", "gray"),legend = "top", 
   ggtheme = ggplot2::theme_minimal())
q<-p+theme(text = element_text(size = 20))
q
ggsave("MAplot_SvC.tiff",q,device="tiff",dpi=300)

p<-ggmaplot(resA, main = expression("MA-Plot remission vs. controls"),
   fdr = 0.05,fc = 0.0, size = 4, top = 0,
   palette = c("#B31B21", "#1465AC", "gray"),
   genenames = as.vector(resS$name),
   legend = "top", 
   font.label = c("bold", 13),
   font.legend = "bold",
   font.main = "bold",
  select.top.method = c("padj","fc"),
   ggtheme = ggplot2::theme_minimal())
q<-p+theme(text = element_text(size = 20))
ggsave("MAplot_AvC.tiff",q,device="tiff",dpi=300)
q

p<-ggmaplot(resSvA, main = expression("MA-Plot exacerbation vs. remission"),
   fdr = 0.05, fc = 0.0,size = 4, top = 0,
   palette = c("#B31B21", "#1465AC", "gray"),
   genenames = as.vector(resS$name),
   legend = "top",
   font.label = c("bold", 13),
   font.legend = "bold",
   font.main = "bold",
   alpha = 0.01,
    select.top.method = c("padj","fc"),
   ggtheme = ggplot2::theme_minimal())
q<-p+theme(text = element_text(size = 20)) 
ggsave("MAplot_SvA.tiff",q,device="tiff",dpi=300)
q

p<-ggplot(as.data.frame(resS), aes(x=pvalue)) + 
  geom_histogram(color="black", fill="grey")+ theme_classic()+
  ggtitle("Exacerbation vs. Controls")
q<-p+ theme(text = element_text(size = 20))
ggsave("pvalhist_SvC.tiff",q,device="tiff",dpi=300)
q

p<-ggplot(as.data.frame(resA), aes(x=pvalue)) + 
  geom_histogram(color="black", fill="grey")+ theme_classic()+
  ggtitle("Remission vs. Controls")
q<-p+ theme(text = element_text(size = 20)) 
ggsave("pvalhist_AvC.tiff",q,device="tiff",dpi=300)
q

p<-ggplot(as.data.frame(resSvA), aes(x=pvalue)) + 
  geom_histogram(color="black", fill="grey")+ theme_classic()+
  ggtitle("Exacerbation vs. Remission")
q<-p+ theme(text = element_text(size = 20)) 
ggsave("pvalhist_SvA.tiff",q,device="tiff",dpi=300)
q
```


```
